---
title: "Class11: Genomics"
author: "Alice (PID:A16799081)"
format: pdf
---

## Section 1. Proportion of G/G in a population 

```{r}
mxl <- read.csv("MXL.csv")
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

Now let's look at GBR: 

```{r}
gbr <- read.csv("GBR.csv")
```

```{r}
table(gbr$Genotype..forward.strand.)
```

```{r}
table(gbr$Genotype..forward.strand.) / nrow(gbr) * 100
```


## Section 4: Population Scale Analysis

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
```

```{r}
nrow(expr)
```

The size of each genotypes are as follows: 

```{r}
table(expr$geno)
```

```{r}
summary(expr)
```

```{r}
medians <- tapply(expr$exp, expr$geno, median)

medians
```

A/A has 108 samples, A/G has 233 samples, and G/G has 121 samples. The corresponding median expression levels are 31.25, 25.06, and 20.07. 


> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
library(ggplot2)

ggplot(expr) + aes(geno, exp, fill=geno) + 
  geom_boxplot(notch = T) 
```

A/A have a higher expression vablue than G/G as displayed in the plot. This infers that the SNP does affect the expression of ORMDL3. 

 




