---
title: "Class8: Breast Cancer Mini Project"
author: "Alice Lai (PID: A16799081)"
format: pdf
---
## About 

In today's lab we will work with fine needle aspiration(FNA) of breast mass data from the University of Wisconsin.  


## Data import  

```{r}
# Save your input data file into your Project directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data, row.names=1)
```

> Q1. How many observations/patients/samples are in this dataset? 

```{r}
obs <- nrow(wisc.df)
```
569 observations are in this dataset. 

> Q2. How many of the observations have a malignant diagnosis? 

```{r}
length(which(wisc.df$diagnosis == 'M'))

sum(wisc.df$diagnosis == 'M')

table(wisc.df$diagnosis)
```
212 observations have a malignant diagnosis. 

> Q3. How many variables/features in the data are suffixed with _mean? 

```{r}
colnames(wisc.df)
```

```{r}
inds <- grep("_mean", colnames(wisc.df), value=T)
length(inds)
```
10 variables are suffixed with _mean. 


## Initial analysis 

Before further analysis, I want to take out the expert diagnosis column (aka the answer) from the dataset 

```{r}
diagnosis <- as.factor(wisc.df$diagnosis)
head(diagnosis)

# We can use -1 here to remove the first column
wisc.data <- wisc.df[,-1]
```


## Clustering 

let's try `kmeans()`

```{r}
km = kmeans(wisc.data, centers = 2)
km$cluster
```

```{r}
table(km$cluster)
```

```{r}
table(km$cluster, diagnosis)
```
let's try `hclust()`

```{r}
hc <- hclust(dist(wisc.data))
```

i can make a tree like figure  

```{r}
plot(hc)
```


## PCA 

Do we need to scale the data? 
we can look at the sd of each column (original variable)

```{r}
round(apply(wisc.data, 2, sd))
```
Yes we need to scale 

```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

Generate out main PCA plot (score plot, PC1 vs PC2 plot)...

```{r}
library(ggplot2)

res <- as.data.frame(wisc.pr$x)
```

```{r}
ggplot(res) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```


## Combining methods 

Clustering on PCA resutls 

```{r}
d <- dist(wisc.pr$x[,1:3])
hc <- hclust(d, method="ward.D2")
plot(hc)
```

To get my clustering result/membership vector I need to "cut" the tree with the `cutree` function 

```{r}
grps <- cutree(hc, k=2)
```

> Q. how many patients in each cluster group? 

```{r}
table(grps)
```

```{r}
plot(res$PC1, res$PC2, col=grps)
```

## Prediction 

we cam use our PCA results to do predictions, that is take new unseen data and project it onto our new PC variables 
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
```

```{r}
plot(res$PC1, res$PC2, col=grps) 
points(npc[,1], npc[,2], col="blue", pch=16, cex=3) 
text(npc[,1], npc[,2], labels=c(1,2), col="white")
```

## Summary 

Principal Component Analysis (PCA) is a super useful method for analyzing larger dataset. It works by finding new variables that capture the most variance from the original variables in your dataset. 
